### 2023-02-09 conclusion.Rmd
Adding first cut of thesis

---
title: ''
output: pdf_document

---

\addcontentsline{toc}{section}{Conclusion}
\section*{Conclusion}

Thank you for following this tutorial!

I hope you'll find it useful to write a very professional dissertation.

\cite{ngTracingPathChikungunya2010}