### 2023-02-09 frontmatter.Rmd

---
title: ''
output: pdf_document
indent: true
---

\addcontentsline{toc}{section}{Acknowledgements}

\section*{Acknowledgements}

<!-- Thank you to: -->

<!-- Stéphane Verguet -->
<!-- Margaret McConnell -->
<!-- Jessica Cohen -->

<!-- Feedback from: -->

<!-- <!-- Sebastian Bauhoff (early ideas) -->
<!-- <!-- Kevin Croke (early ideas) --> 
<!-- <!-- Harvard Center for Population and Development Studies  --> 
<!-- <!-- David Canning (teaching and econometrics) -->

<!-- Support from: -->

<!-- Jarvis Chen -->
<!-- Bruce Villineau -->
<!-- Matthew Boccuzzi -->
<!-- Claudette Agustin -->

<!-- Allison Conary -->
<!-- Barbara Heil -->

<!-- Friends at HSPH: -->
<!-- ... -->
<!-- Family: -->
<!-- Partner: -->

\newpage

\hspace{0pt}

\vfill

\begin{raggedright}

\emph{Dedicated to ....}

\end{raggedright}

\vfill

\hspace{0pt}