### 2023-02-09 paper2.Rmd

---
title: ''
output: pdf_document

---

\addcontentsline{toc}{section}{Paper 2}
\section*{Paper 2}

Thank you for following this tutorial!

I hope you'll find it useful to write a very professional dissertation.

\cite{ngTracingPathChikungunya2010}