### 2023-02-22 appendix.Rmd
Addition of appendices, revisions throughout papers, acknowledgements

\subsection*{Appendix 1}

\addcontentsline{toc}{subsection}{Appendix 1}

```{r appendix_1, child='appendix_1.Rmd'}
```

\newpage

\subsection*{Appendix 2}

```{r library_appendix, warning = FALSE, message = FALSE, echo = FALSE, include = FALSE}

# Loading libraries
# library(MASS) # We load this and then detach it just to clear the package
# library(margins) # We load this and then detach it just to clear the package
# library(lme4) # We load this and then detach it just to clear the package
# detach("package:lme4", unload=TRUE) # Needed to use dplyr's select function
# detach("package:margins", unload=TRUE) # Needed to use dplyr's select function
# detach("package:MASS", unload=TRUE) # Needed to use dplyr's select function
library(readxl)
library(readr)
library(dplyr)
library(zoo)
library(tidyr)
library(psych)
library(ggplot2)
library(RColorBrewer)
library(sp)
library(broom)
library(sf)
library(scales)
library(viridis)
library(data.table)
library(ggpubr)
library(ggmap)
library(maps)
library(mapdata)
library(ggrepel)
library(table1)
library(stringr) 
library(stringi)
library(fixest)
library(stats)
library(lubridate)
library(forecast)
library(gtools)
# library(devtools)
library(reticulate)
library(stargazer)
library(lwgeom)
if (!require("ggsflabel")) devtools::install_github("yutannihilation/ggsflabel")
library(openxlsx)
library(gganimate)
library(transformr)
library(huxtable)
library(ggtext)
library(margins)
library(ncdf4) # package for netcdf manipulation
library(magrittr)
library(lmtest)
library(sandwich)
library(rgee)
library(gt)
library(knitr)
library(webshot2)
library(sysfonts)
library(showtext)
font_add_google("Source Sans Pro")
library(gridExtra)
library(tibble)
showtext_auto() 
showtext_opts(dpi = 300)  
# library(googledrive)
# drive_auth(email = TRUE)
library(rmdHelpers)
library(gtExtras)
library(kableExtra)
library(DescTools)
library(rvest)
library(httr)
library(polite)
library(janitor)
library(cowplot)
library(ggridges)
# library(ggsn)

# library(glmmTMB)

# Setting directories - change these to your local directory
path <- 'C:/Users/danie/Dropbox (Harvard University)/madagascar_mh'
code <- paste(path,'code', sep = "/")
path_python  = "C:/Users/danie/AppData/Local/Programs/Python/Python310/python.exe"

# Data
datasource <- paste('C:/Users/danie/Dropbox (Harvard University)/data',
                    sep = "/") # The complete data set is not uploaded to 
# Github. Ask golden@hsph.harvard.edu for access. 
datapath <- paste(path,'data', sep = "/")
datapath_reference  <- paste(datapath,'0_reference', sep = "/")
data_mh_chd  <- paste(datapath,'1_mental_health', sep = "/")
data_mh_csb <- data_mh_chd  
data_mh_chu_chrr <- paste(data_mh_chd,'pathologies neuropsychiatriques CHU CHRR', sep = "/")
data_gps <- paste(datapath,'2_gps', sep = "/")
data_gps_sources <- paste(data_gps,'gps_sources', sep = "/")
data_gps_sources_searches <- paste(data_gps_sources,'searches', sep = "/")
data_gps_sources_reposit <- paste(data_gps_sources,'repositories', sep = "/")
data_climate_path  <- paste(datapath,'3_climate', sep = "/")
data_climate_era  <- paste(data_climate_path,'ERA5', sep = "/")
data_climate_esa  <- paste(data_climate_path,'ESA_CCI', sep = "/")

# Results
resultspath <- paste(path,'results', sep = "/")
mh_path <- paste(resultspath,'1_mental_health', sep = "/")
gpspath <- paste(resultspath,'2_gps', sep = "/")
gpspath_admin_names <- paste(gpspath,'0_reference', sep = "/")
gpspath_fig <-  paste(gpspath,'1_figures', sep = "/")
gpspath_spread <-  paste(gpspath,'2_spreadsheets', sep = "/")
gpspath_spread_fuzzy <-  paste(gpspath_spread,'3_fuzzy_matching', sep = "/")
analysispath <- paste(resultspath,'4_analysis', sep = "/")
analysis_figures <- paste(analysispath,'1_figures', sep = "/")
analysis_models <- paste(analysispath,'3_models', sep = "/")

# Shapefiles
mappath <- paste(path,'shapefiles', sep = "/")
mappath_MOH <- paste(mappath,'MOH', sep = "/")
mappath_OCHA <- paste(mappath,'OCHA', sep = "/")

# Color palette 
col_light = "cyan4"
col_dark = "cyan3"
col_bright = "coral"
col_black = "black"
col_grey_dark = "grey20"
col_grey_med = "grey40"
col_grey_light = "grey80"
col_grey_lightest = "grey85"
col_soilmoisture = col_dark
col_temperature = col_light
col_interaction = col_bright
col_cyclone = "steelblue2"

pal <- c("lightskyblue2", "steelblue2", "dodgerblue2", "dodgerblue3", "dodgerblue4", "darkblue")

col_within_threshold = "cyan3"
col_beyond_threshold = "coral"

col_centroid <- "grey20"
col_exact_all <- "cyan2"
col_exact_rhino <- "cyan3"
col_exact_val <- "cyan4"
col_search <- "coral"

col_CH <- "coral"
col_CHD <- "coral1"
col_CHR <- "coral3"
col_CSB <- "steelblue2"
col_CSB1 <- "steelblue1"
col_CSB2 <- "steelblue3"
col_other <-  "grey20"

col_model_main <- "#F3AB00"
col_model_other <- col_grey_med
col_spec_yes <-  "#73BDBF"
col_spec_no <-  "white"

col_cases <- "chartreuse3"
col_no_cases <- "deeppink3"

col_eco_rainforest <-"#32A42F"
col_eco_highlands <- "#B9CF19"
col_eco_arid <- "#FEEE19"
col_eco_drydecid <- "#FCAF3E"

# Thresholds
# These are cut off values that will be used in the GPS do file

length_lim <- 15  # The maximum distance allowed when the same facility has
# two GPS coordinates in the data sources for the center of
# those two coordinates to be used as the facility's location.
# Note: the selection of 15 is arbitrarily conservative. Spatial resolution of 
# climate data is typically ~ 30km x 30 km

area_lim <- length_lim*length_lim

# The maximum area of the convex hull allowed when the same facility has
# 3+ GPS coordinates in the data sources for the centroid of
# that convex hull to be used as the facility's location

fokontany_lim <- area_lim    # A small fokotanty is one that is ____ km ^2 or smaller

commune_lim <- area_lim    # A small commune is one that is ____ km ^2 or smaller

# Lags: specify how many lags you want for the explanatory variables (e.g., soil moisture, temp)

lags_number <- 18

# Climate data start and end date
start_date <- as.Date("2010-01-01")
end_date <- as.Date("2020-07-01")

start_date_climate <- start_date - lags_number*31 #This ensures there is data for the lagged exposure
end_date_climate <- end_date

# ESA CCI trim dates
date1 <- 200801
date2 <- 202112

# Seed and sampling fraction
set.seed(1960) # Madgascar's year of independence
sample_fraction = .01

######################
##      MAPS        ##
######################
#   
# # There are two map options available. One set of shapefiles were available
# # from humdata.org, while the other set was provided by the MOH.
#   
# # Loading high-resolution maps of St. Lucia (from data.humdata.org)
map_mdg_0_un <- st_read(paste0(mappath_OCHA,"/",
                               "mdg_admbnda_adm0_BNGRC_OCHA_20181031.shp"))
map_mdg_1_un <- st_read(paste0(mappath_OCHA,"/",
                               "mdg_admbnda_adm1_BNGRC_OCHA_20181031.shp"))
map_mdg_2_un <- st_read(paste0(mappath_OCHA,"/",
                               "mdg_admbnda_adm2_BNGRC_OCHA_20181031.shp"))
map_mdg_3_un <- st_read(paste0(mappath_OCHA,"/",
                               "mdg_admbnda_adm3_BNGRC_OCHA_20181031.shp"))
loaded_maps_un <- 1

# Load in data from MOH
map_mdg_0 <- st_read(paste0(mappath_MOH,"/",
                            "adm0_mada_utm38S.shp"))
map_mdg_1 <- st_read(paste0(mappath_MOH,"/",
                            "adm1_mada_utm38S.shp"))
map_mdg_2 <- st_read(paste0(mappath_MOH,"/",
                            "adm2_mada_utm38S.shp"))
map_mdg_3 <- st_read(paste0(mappath_MOH,"/",
                            "adm3_mada_utm38S.shp"))
map_mdg_4 <- st_read(paste0(mappath_MOH,"/",
                            "adm4_mada_utm38S.shp"))

```

\addcontentsline{toc}{subsection}{Appendix 2}

\subsubsection*{Appendix 2.1: Health system data cleaning}

\addcontentsline{toc}{subsubsection}{Appendix 2.1}

```{r appendix_2, child='appendix_2.1.Rmd'}
```
\newpage

\subsubsection*{Appendix 2.2: Health facility geocoding and validation}

\addcontentsline{toc}{subsubsection}{Appendix 2.2}

```{r appendix_3, child='appendix_2.2.Rmd'}
```

\newpage