### 2023-02-25 conclusion.Rmd
Abstract and Introduction for Paper 3

\section*{Conclusion}

\addcontentsline{toc}{section}{Conclusion}

Overview of Paper 1

Overview of Paper 2

Overview of Paper 3

Policy implications and future research

Burden may be larger than traditionally conceptualized

Papers also show that mental health is linked to broader contextual challenges (climate change, COVID-19) – resilience/vulnerability, exacerbating existing disparities, but also could be critical to overcoming them…

In the absence of population mental health metrics, innovative uses of health data may be helpful in conducting research, but come with limitations