### 2023-02-25 intro.Rmd

\section*{Introduction}

\addcontentsline{toc}{section}{Introduction}

Overview of the global burden of mental health

Cite key literature – systematic reviews, Lancet series, Vikram’s work, DC3P findings, Walker et al. EMERALD health systems & mental health study, Mental Health Atlas from WHO, GBD for estimates of burden

Implications of this global burden for health systems

Highlight interconnectedness of mental health and physical health

Links of mental health to financial risk protection / vulnerabilities to poverty

Underscore under-investment in mental health systems, coupled with lack of data on mental health, which constrains research and policy

Also note that mental health is a truly global issue, impacting HICs and LMICs alike

Overview of Paper 1

Overview of Paper 2

Overview of Paper 3