### 2023-03-11 conclusion.Rmd

\section*{Conclusion}

\addcontentsline{toc}{section}{Conclusion}

In this dissertation, I investigated different quantitative facets of the global burden of mental disorders, a burden that is widespread,[@charlsonExcessMortalityMental2015; @rehmGlobalBurdenDisease2019] growing,[@patelLancetCommissionGlobal2018] and largely unmet.[@bassUnitedNationsGeneral2012a; @demyttenaerePrevalenceSeverityUnmet2004]

In Chapter 1, we examined how the global burden is defined, and we proposed an estimation approach to more comprehensively reflect morbidity and loss of life due to suicide and premature mortality attributable to mental disorders. After applying this composite estimation approach to the most recently available estimates from the Global Burden of Disease study, we found that approximately one in 6 disability-adjusted life years (approximately 16%) were due to mental disorders in 2019. We then applied a value of a statistical life year approach to estimate the associated global economic burden. Our findings suggest staggering loses: after accounting for premature mortality and other sources of morbidity associated with mental disorders, global losses were estimated to approximate 4.7 trillion USD in 2019. The magnitude of these findings underscores the global scale of impact that mental disorders have and highlights the urgency for research and action to support mental health worldwide.

In Chapter 2, we examined how this global burden might be shifting with changes in climate by analyzing health system data in Madagascar, a country with little to no mental health surveillance data and high vulnerability to climate change. We obtained data on monthly counts of mental health related visits reported by 3,413 facilities in Madagascar from 2010 to 2020. We then geolocated these facilities and obtained meteorologic data on climate conditions to examine the impact of three climate change exposures—temperature, soil moisture, and duration of cyclones and tropical storms. We found that monthly mental health related visits declined as temperature increased, particularly in the central highlands, where cooler temperatures were more prevalent; conversely, higher grades of soil moisture were associated with greater mental health visits with a lag of approximately three months, an increase found primarily in regions associated with higher flood risk in Madagascar. Our findings illustrate potential heterogeneity in how changing climate conditions may impact mental health needs, and our work demonstrates how routine health system data may be leveraged to address data gaps in health surveillance, particularly in resource-constrained settings.

In Chapter 3, I examined how search data for health symptoms could, in the absence of high frequency surveillance data, provide insight into changes in population mental health during the COVID-19 pandemic. I obtained Google search data from U.S. counties, states, and international trends for over 400 symptoms and used these data to study the immediate and sustained impacts of announcements of the development of safe and effective COVID-19 vaccines in November 2020. I found that searches for anxiety, depression, and major depressive disorder significantly fell after these announcements, and that this decline continued throughout the subsequent weeks. Furthermore, despite the subsequent (and ongoing) politicization of COVID-19 vaccines in the United States, I observed consistency in search density declines for mental health symptoms across red and blue states/counties. The findings speak to the importance of timely and transparent communication in public health, as well as to the potential use of internet search data to investigate changes in population mental health that may be difficult to detect without high frequency data.

These papers add to a growing understanding of the tremendous scale of mental health needs worldwide. The magnitude of this burden has galvanized a global movement and a call to action for greater investment and prioritization for mental health.[@patelRenewedAgendaGlobal2011] This movement has emphasized the importance of investing in mental health as a means of promoting sustainable development, human rights, and social inclusion.[@patelLancetCommissionGlobal2018]

To respond to this burden, health systems will need to be mindful of factors beyond the traditional extent of mental health services and consider mental health within a background of broader contextual challenges. As seen in Chapters 2 and 3, changes in climate and the course of the COVID-19 pandemic highlight the importance of considering the connections between mental health and other issues in public health. Insofar as these challenges may exacerbate one another and strengthen vulnerabilities and inequities in health, responding to them in concert—and addressing the mental health needs of populations facing new and evolving stressors—may be key in overcoming them and fostering resiliency. This will require global, multisectoral action across countries at all economic levels, coupled with a substantive increase in financial resources dedicated to mental health.

Moreover, health systems and global health research will need to develop new means of dealing with the current limitations in the measurement of population mental health. As discussed in Chapter 1, traditional approaches to estimating the global burden of mental disorders may not capture its true scale, contributing to a lack of perceived importance of mental health to human well-being. This perception remains unchallenged, in part, due to the lack of detailed epidemiological data on the prevalence and distribution of mental disorders, a challenge that further constrains mental health research. As demonstrated by Chapters 2 and 3, innovative uses of health system and internet search data may help in overcoming this constraint but come with limitations of their own.

While this dissertation has underscored that the burden of mental disorders is complex in nature and vast in scope, the challenges this burden poses are not insurmountable. By measuring and understanding the mental health needs of populations, and by addressing these needs within a broader context of societal challenges, health systems may more effectively respond to this challenge, bringing the world closer to parity between mental health and physical health. Through a global effort to address this essential facet of human wellbeing, millions of individuals may be helped to achieve their full potential, strengthen resilience, and obtain access to the care they need to flourish.