### 2023-05-11 thesis_final_version.Rmd

---
title: " "		
output:
  pdf_document:
    pandoc_args: !expr acronymsdown::add_filter()
    number_sections: TRUE

  # pdf_document:
  #   pandoc_args: !expr acronymsdown::add_filter()
  #   number_sections: TRUE

reference-section-title: Works Cited

geometry: margin=1in
fontsize: 11pt
indent: true
header-includes:
  - \usepackage{sectsty}
  - \usepackage{setspace}\spacing{2}
  - \usepackage{fancyhdr}
  - \usepackage[utf8]{inputenc}
  - \usepackage[utf8]{lscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
  - \usepackage[nottoc]{tocbibind}
  # - \usepackage{acro}
  # - \usepackage{float}
  # - \usepackage{chapterbib}
  # - \usepackage{lastpage}
  # - \usepackage{dcolumn}
  # - \usepackage[backend=biber,style=authoryear-icomp]{biblatex}
  # - \addbibresource{bibliography.bib}

bibliography: main.bib
csl: https://www.zotero.org/styles/the-lancet
acronyms:
  insert_loa: false
  insert_links: false
  loa_title: ''
  keys:
    - shortname: DALYs
      longname: disability-adjusted life years
    - shortname: YLLs
      longname: years of life lost 
    - shortname: YLDs
      longname: years of healthy life lost due to disability 
    - shortname: LMICs
      longname: low- and middle-income countries
    - shortname: WHO
      longname: World Health Organization
    - shortname: USD
      longname: United States dollars
    - shortname: GBD
      longname: Global Burden of Disease
    - shortname: GBO
      longname: Global Health Observatory
    - shortname: GDP
      longname: gross domestic product
    - shortname: NCDs
      longname: noncommunicable diseases
    - shortname: PAF
      longname: population attributable fraction
    - shortname: VSL
      longname: value of a statistical life
    - shortname: IRB
      longname: Institutional Review Board 
    - shortname: UIs
      longname: uncertainty intervals
    - shortname: CIs
      longname: confidence intervals
    - shortname: PPP
      longname: purchasing power parity
    - shortname: CCEs
      longname: climate change-related exposures
    - shortname: HMIS
      longname: health management information system
    - shortname: ESA CCI
      longname: European Space Agency’s Climate Change Initiative
    - shortname: ESA CCI SM
      longname: ESA CCI soil moisture dataset
    - shortname: ERA5
      longname: 'European ReAnalysis 5' 
    - shortname: NOAA
      longname: United States National Oceanic and Atmospheric Administration
    - shortname: GESIS
      longname: Gestion du Système d'Information Sanitaire
    - shortname: DHIS2
      longname: 'District Health Information Software 2'
    - shortname: CSB
      longname: centre santé de base
    - shortname: FSB
      longname: formation sanitaires privées de dase
    - shortname: CHD/CHRD
      longname: centre hospitalier de référence de district
    - shortname: RHINO
      longname: Routine Health Information Network
    - shortname: USAID
      longname: United States Agency for International Development
    - shortname: US
      longname: United States
    - shortname: CHRR
      longname: centre hospitalier de référence régionale
    - shortname: CHU
      longname: centre hospitalier universitaire
    - shortname: C3S
      longname: Copernicus Climate Change Service
    - shortname: ECMWF
      longname: European Centre for Medium-Range Weather Forecasts
    - shortname: km
      longname: kilometers
    - shortname: MFR La Réunion
      longname: Météo-France La Réunion
    - shortname: WMO
      longname: World Meteorological Organization
    - shortname: RSMC 
      longname: Regional Specialized Meteorological Center
    - shortname: NB
      longname: negative binomial
    - shortname: IRRs
      longname: incidence rate ratios
    - shortname: FE
      longname: fixed effects
    - shortname: CDC
      longname: Centers for Disease Control and Prevention
    - shortname: FDA
      longname: Food and Drug Administration
    - shortname: EUA
      longname: Emergency Use Authorization
    - shortname: MIT
      longname: Massachusetts Institute of Technology
    - shortname: FIPS
      longname: Federal Information Processing Standards
    - shortname: USDA
      longname: United States Department of Agriculture
    - shortname: ITS
      longname: interrupted time series
    - shortname: CITS
      longname: controlled interrupted time series
    - shortname: MDD
      longname: major depressive disorder
    - shortname: ADHD
      longname: attention deficit hyperactivity disorder
    - shortname: CoC
      longname: clouding of consciousness
    - shortname: GAD
      longname: generalized anxiety disorder
---

```{r setup0, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, include = FALSE, echo = TRUE) 
library(acronymsdown)
```

\allsectionsfont{\centering\normalsize}
\subsectionfont{\raggedright}
\subsubsectionfont{\raggedright}

\pagenumbering{roman}
\pagestyle{empty}

\begin{centering}

\vspace*{36pt}

\large

{\bf `r toupper("Quantitative Approaches Towards Evaluating the Global Burden of Mental Illness")`}

\vspace{12pt}

\normalsize

by

\vspace{12pt}

{\ Daniel Alejandro Arias}

\vspace{36pt}

A dissertation submitted to the faculty of \\
the Harvard T.H. Chan School of Public Health\\
in partial fulfillment of the requirements\\
for the degree of \emph{Doctor of Philosophy}\\
in the Department of Global Health and Population

\vspace{42pt}

Harvard University \\
Cambridge, Massachusetts

\vspace{42pt}

`r format(Sys.time(), '%B %Y')`

\end{centering}

\newpage

\vspace*{\fill}
© `r format(Sys.time(), '%Y')` Daniel Arias. All rights reserved.

\newpage

\pagestyle{plain}

\normalsize

\begin{centering}

Dissertation Advisor: Stéphane Verguet \hfill Daniel Arias

\vspace{24 pt}

{\bf `r toupper("Quantitative Approaches Towards Evaluating the Global Burden of Mental Illness")`}

\vspace{12pt}

\section*{Abstract}

\addcontentsline{toc}{section}{Abstract}

\end{centering}

Mental disorders are a leading cause of morbidity and mortality, affecting over one billion people worldwide. This burden, across all levels of socioeconomic development, is rising, yet the majority of people living with a mental disorder do not have access to adequate mental health services, with many facing stigma and high out-of-pocket payments for their care. Moreover, new and emerging dangers such as climate change and the COVID-19 pandemic pose additional stressors that threaten to exacerbate this burden.

This dissertation aims to contribute to a better understanding of the evolving burden and dynamics of global mental health. The three studies it presents take as their focus different quantitative aspects of how the global burden is measured and evaluated, how it may be changing in response to global challenges, and how novel data sources may help in characterizing these changes, particularly in the absence of robust, high-frequency epidemiological surveillance.

In Chapter 1, we investigated the share of the global burden of disease attributable to mental disorders and its associated economic value. To capture premature mortality due to mental disorders, as well as disability from associated causes, we proposed a composite approach to estimation. Using the most recently available estimates from the Global Burden of Disease study, we found that the burden of mental disorders is likely much higher than previously estimated, encompassing 16% of disability-adjusted life years in 2019. The economic value of this mental health burden was estimated to exceed 4.7 trillion United States dollars using a value of a statistical life year approach, accounting for regional losses that range from 3.9% of gross domestic product in Eastern Sub-Saharan Africa to 7.9% in High-income North America.

In Chapter 2, we focused on Madagascar—a low-income country that is among the most vulnerable to the health consequences of climate change—and examined how health system visits for mental disorders may be shifting in response to climate change exposures such as changes in temperature, soil moisture, and the duration of tropical storms and cyclones. Drawing on meteorologic, geospatial, and health system data reported by 3,413 facilities from 2010 to 2020, we conducted an ecological analysis using negative binomial regression. Our results indicated that warmer temperatures in the cooler central highlands were associated with a decrease in monthly reported visits for mental disorders, while higher soil moisture could lead to an increase in visits, particular in high flood-risk regions and after a three-month lag, indicating potential variation in the impact of climate change on mental health needs and system responses.

In Chapter 3, I investigated how Google search data for mental health symptoms might provide insight into the state of mental health during the COVID-19 pandemic. Using an interrupted time series approach with search data from Australia, Ireland, New Zealand, Singapore, the United Kingdom, and the United States, I found that announcements of COVID-19 vaccine safety and efficacy data in November 2020 were associated with immediate and sustained declines in search density for anxiety and depression. These declines in searches, if taken as a reasonable proxy for population mental health, underscore the importance of timely and transparent public health communication and illustrate the potential application of high-frequency internet search data for population mental health surveillance.

Taken together, these chapters highlight different facets of the global burden of mental disorder and contribute to a growing effort to generate a more comprehensive understanding of the challenges posed by this burden at local, regional, and global levels.

\cleardoublepage

\begin{spacing}{1}

\renewcommand*\contentsname{Table of Contents}

\tableofcontents

\addcontentsline{toc}{section}{Table of Contents}

\newpage

\listoffigures

\newpage

\listoftables

\newpage

\cleardoublepage

\newpage

\end{spacing}

\section*{List of Acronyms}

\addcontentsline{toc}{section}{List of Acronyms}

\printacronyms

\newpage

```{r frontmatter, child='frontmatter.Rmd'}
```

\cleardoublepage

\pagenumbering{arabic}

```{r intro, child='intro.Rmd'}
```

\cleardoublepage

```{r paper1, child='paper1.Rmd'}
```

\cleardoublepage

```{r paper2, child='paper2.Rmd'}
```

\cleardoublepage

```{r paper3, child='paper3.Rmd'}
```

\cleardoublepage

```{r conclusion, child='conclusion.Rmd'}
```

\cleardoublepage

\section*{References}

\addcontentsline{toc}{section}{References}

<div id="refs"></div>

\cleardoublepage

\section*{Supplementary materials}

\addcontentsline{toc}{section}{Supplementary materials}

```{r appendix_all, child='appendix.Rmd'}
```

\newpage

<!-- ```{r todo} -->

<!-- ### TO DO / KNOWN ISSUES -->

<!-- # Check that sequential articles are represented by a dash -->
<!-- # Institutional authors need to be fixed in bibliography -->
<!-- # Numbered figures and tables by section -->

<!-- ``` -->